---
title: "p8105_hw6_mnt2130"
author: "Mindy Tran"
date: "2022-12-03"
output: github_document 
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(patchwork)
library(dplyr)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

This code below obtains a distribution for $\hat{r}^2$ by drawing bootstrap samples; the a model to each andextracts the value I'm concerned with and summarizes it summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r weather_df, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


```{r}
weather_df %>% 
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density() +
   labs(
    x = "R Squared Values",
    y = "Density",
    title = "Distribution of R-Squared estimates for Central Park Weather"
  ) 

```

Based on the results, the $\hat{r}^2$ value is high. The upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for $R^2$, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, since the shape of the curve generated isn't symmetric, using the mean +/- 1.96 times the standard error would probably not work well in this case. 

This code produces a distribution for $\log(\beta_0 * \beta1)$ using a similar approach. 

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density() +
  labs(
    x = "Log(B0*B1)",
    y = "Density",
    title = "Distribution of Log(B0*B1) estimates for Central Park Weather"
  ) 
```

As with $r^2$, this distribution is approximately normally distributed but also somewhat skewed and potentially has some outliers. 

### Problem 2
The following code chunk creates a city_state variable and a binary variable which indicates whether a homicide is resolved. It also omits the cities : Tulsa, Dallas, Phoenix, and Kansas which have missing data or data entry mistakes. We also limited our analyses to victims who were white or black and converted the variable victim_age to a numeric variable.

```{r homicide}

homicide_df = 
  read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = "_"),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1,
    ),
    victim_age = as.numeric(victim_age)
  ) %>% 
  filter(!city_state %in% c("Tulsa_AL", "Dallas_TX", "Phoenix_AZ","Kansas City_MO"))  %>% 
  filter(victim_race %in% c("Black", "White")) %>% 
  filter(victim_sex %in% c("Male", "Female")) %>%  
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```
```{r baltimore}
baltimore_logitstic = homicide_df %>% 
  filter(city_state == "Baltimore_MD") %>% 
glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial() , data = .) %>%
  broom::tidy() %>% 
  mutate(
    odds_ratio = exp(estimate),
    ci_lower =  exp(estimate - 1.96 * std.error),
    ci_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, odds_ratio, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
baltimore_logitstic
```

